### 2018-06-17 CS761-RandomizedAlgorithms/assignments/a2.tex
progress

\paragraph{}
Suppose we have a list $L$ of $n$ distinct numbers, whose sorted list would we $\{x_1,\dots, x_n\}$. We want to find a $k$-approximate median. I.e. we want to return some $x \in [x_{\frac{n}{2} - k}, x_{\frac{n}{2} + k}]\cap L$. Consider sampling $m$ numbers from $L$ and returning the median $c$ of these $m$ numbers as the true median of $L$.
\paragraph{}
Let $L'$ be list of $m$ sampled numbers. For each $i \in L'$ let $X_{i,<}$ be an indicator variable for the event that $i \in \{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then, since we sample without replacement,
$$Pr(X_{i, <} =1) = (\frac{n}{2} -k -1) / n = \frac{n-2k-2}{2n}$$
Let $X_<$ be the number of elements in $L'$ which lie in $\{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then $X_< = \sum_{i \in L'} X_{i,<}$.
Let $X_>$ be the number of elements in $L'$ which lie in $\{x_{\frac{n}{2} + k + 1}, \dots, x_n\}$. Since we sample uniformly and
	$$|\{x_{\frac{n}{2} + k + 1}e, \dots, x_n\}| =  |\{x_1, \dots, x_{\frac{n}{2} -k -1}\}| = \frac{n}{2} - k - 1$$
we observe that $X_<$ and $X_>$ have the same distribution.
\paragraph{}
Now we see that $c$ fails to be a $k$-approximate median for $L$ if and only if $X_< \geq \frac{m}{2}$ or $X_> \geq \frac{m}{2}$. Since $X_<$ and $X_>$ have the same distribution, the probability of failure is thus
$$2Pr(X_< \geq \frac{m}{2}).$$
We plan to bound this using a Chernoff bound. First we compute the expectation,
$$E[X_<] = \sum_{i\in L'} Pr(X_{i,<} = 1)  = \sum_{i \in L'} \frac{n-2k-2}{2n} = \frac{m(n-2(k+1))}{2n}.$$
So 
$$Pr(X_< \geq \frac{m}{2}) = Pr(X_< \geq (1+\delta) E[X_<])$$
where $$\delta =\frac{2(k+1)}{n-2(k+1)}.$$
We proceed by assuming $k < \frac{n}{4}-1$ so that $0<\delta <1$. Ultimately we are interested in results on how many samples $m$ we need to take to obtain a certain quality of approximation, and so bounding $k$ is reasonable.
\paragraph{}
Since $0 < \delta < 1$, and $X_<$ is a sum of independent random coin flips, we can apply the second Chernoff bound from Theorem on page $4$ of Lecture $2$, obtaining
$$Pr(X_< \geq \frac{m}{2}) < \exp(\frac{-\delta^2E[X_<]}{3}).$$
%Problem 4
\newpage
\section{}

%Problem 5
\newpage
\section{}
\end{document}

>>>>

\paragraph{}
Suppose we have a list $L$ of $n$ distinct numbers, whose sorted list would we $\{x_1,\dots, x_n\}$. We want to find a $k$-approximate median. I.e. we want to return some $x \in [x_{\frac{n}{2} - k}, x_{\frac{n}{2} + k}]\cap L$. Consider sampling $m$ numbers from $L$ and returning the median $c$ of these $m$ numbers as the true median of $L$.
\paragraph{}
Let $L'$ be list of $m$ sampled numbers. For each $i \in L'$ let $X_{i,<}$ be an indicator variable for the event that $i \in \{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then, since we sample without replacement,
$$Pr(X_{i, <} =1) = (\frac{n}{2} -k -1) / n = \frac{n-2k-2}{2n}$$
Let $X_<$ be the number of elements in $L'$ which lie in $\{x_1, \dots, x_{\frac{n}{2} -k -1}\}$. Then $X_< = \sum_{i \in L'} X_{i,<}$.
Let $X_>$ be the number of elements in $L'$ which lie in $\{x_{\frac{n}{2} + k + 1}, \dots, x_n\}$. Since we sample uniformly and
	$$|\{x_{\frac{n}{2} + k + 1}e, \dots, x_n\}| =  |\{x_1, \dots, x_{\frac{n}{2} -k -1}\}| = \frac{n}{2} - k - 1$$
we observe that $X_<$ and $X_>$ have the same distribution.
\paragraph{}
Now we see that $c$ fails to be a $k$-approximate median for $L$ if and only if $X_< \geq \frac{m}{2}$ or $X_> \geq \frac{m}{2}$. Since $X_<$ and $X_>$ have the same distribution, the probability of failure is thus
$$2Pr(X_< \geq \frac{m}{2}).$$
We plan to bound this using a Chernoff bound. First we compute the expectation,
$$E[X_<] = \sum_{i\in L'} Pr(X_{i,<} = 1)  = \sum_{i \in L'} \frac{n-2k-2}{2n} = \frac{m(n-2(k+1))}{2n}.$$
So 
$$Pr(X_< \geq \frac{m}{2}) = Pr(X_< \geq (1+\delta) E[X_<])$$
where $$\delta =\frac{2(k+1)}{n-2(k+1)}.$$
We proceed by assuming $k < \frac{n}{4}-1$ so that $0<\delta <1$. Ultimately we are interested in results on how many samples $m$ we need to take to obtain a certain quality of approximation, and so bounding $k$ is reasonable.
\paragraph{}
Since $0 < \delta < 1$, and $X_<$ is a sum of independent random coin flips, we can apply the second Chernoff bound from Theorem on page $4$ of Lecture $2$, obtaining
$$Pr(X_< \geq \frac{m}{2}) < \exp(\frac{-\delta^2E[X_<]}{3}) = \exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))}).$$
Hence the probability of failure is at most
$$2\exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))})$$
To succeed with probability at least $0.9999$ we want this value to be at most $10^{-5}$. So it will suffice to choose $m$ and $k$ to satisfy
$$\exp(\frac{-2(k+1)^2m}{3n(n-2(k+1))}) \leq 10^{-6}$$
Rearranging yields the expression
$$(k+1)^2m \geq 9\ln(10) n^2 - 18\ln(10) n(k+1)$$
Dropping the non-positive term on the left-hand side, as this can only improve our probability of success, gives us the following expression for the trade-off between $m$ and $k$:
$$(k+1)^2m = 9\ln(10)n^2 = \Omega(n^2).$$
\paragraph{}
If we want $k \leq \epsilon n$ for small $\epsilon$ (say $<\frac{1}{4}$) then we need $m$ to satisfy
$$(\epsilon n+ 1)^2 m = 9\ln(10)n^2$$
i.e. 
$$m = \frac{9\ln(10)n^2}{(\epsilon n + 1)^2}.$$
\paragraph{}
Similarly if we want $k \leq n^{1-\eta}$ for $\eta \leq \frac{1}{2}$ then we take (again assuming desired $k$ is less than $\frac{1}{4}n-1$)
$$m =  \frac{9\ln(10)n^2}{(n^{1-\eta} + 1)^2}.$$
$\blacksquare$
%Problem 4
\newpage
\section{}
\paragraph{}
We assume that the determinant of an $n \times n$ matrix where each element is an entry of $\F[x]$ of degree at most $d$ can be computed in $O(dn^\omega)$ field operations.
\subsection{a}
Let $G=(U\cup V, E)$ be a bipartite graph where each edge is coloured either red or blue. We may assume $|V| = |U|$ by adding dummy vertices as needed. Let $k \in \Z$ be such that $1 \leq k \leq |E|$. We construct an $n\times n$ matrix $A$ of indeterminants as follows: for every edge $uv \in E$ set
$$A_{uv} = \begin{cases}
x_{uv}, \text{if $uv$ is coloured blue} \\
rx_{uv}, \text{if $uv$ is coloured red}
\end{cases}$$
with indeterminants $x_{uv}$ for each edge $uv$ and indeterminant $r$.
%Problem 5
\newpage
\section{}
\end{document}
